---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


### Loading and preprocessing the data

```{r}
library(lubridate)
unzip("activity.zip")
data <- read.csv("activity.csv", stringsAsFactors = FALSE)
data$date <- ymd(data$date)
```

### What is mean total number of steps taken per day?

I calculate the total number of steps taken per day

```{r results = 'asis'}
library(xtable)
library(dplyr)
total <- data %>% group_by(date) %>% summarise(sum(steps, na.rm = TRUE))
names(total)[2] <- "steps.per.day"
total$date <- as.character(total$date)
xt <- xtable(total)
print(xt, type = "html")
```

Create a histogram of the total number of steps taken each day. 

```{r}
hist(total$steps.per.day, xlab = "Steps per day", main = "Total steps")
```

Mean of the total number of steps taken per day: 

```{r}
mean_step <- mean(total$steps.per.day)
mean_step
```

Median: 

```{r}
median_step <- median(total$steps.per.day)
median_step
```

## What is the average daily activity pattern?

A time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
steps_by_interval <- data %>% group_by(interval) %>% summarise(mean(steps, na.rm = TRUE))
names(steps_by_interval)[2] <- "average.steps"
plot(steps_by_interval$interval, steps_by_interval$average.steps, type = "l", xlab = "Interval", ylab = "Average number of steps")
```

5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r results='asis'}
maxint <- max(steps_by_interval$average.steps)
number_of_interval <- steps_by_interval %>% filter(average.steps == maxint) 
xt <- xtable(number_of_interval)
print(xt, type = "html")
```

## Imputing missing values

Total number of missing values in the dataset

```{r}
comple <- complete.cases(data)
comple <- comple[comple == FALSE]
length(comple)
```

Fill in all of the missing values for steps in the dataset with the mean for the appropriate 5-minute interval. And create a new data frame with the missing data filled in. Data frame named data1.

```{r}
data1 <- data
for (i in 1:17568) {if (is.na(data1$steps[i])) {                
        inter <- steps_by_interval %>% filter (interval == data1$interval[i])
                data1$steps[i] <- inter$average.steps }
}
```

Create a histogram of the total number of steps taken each day.

```{r}
total1 <- data1 %>% group_by(date) %>% summarise(sum(steps, na.rm = TRUE))
names(total1)[2] <- "steps.per.day"

hist(total1$steps.per.day, xlab = "Total steps per day", main = "Total steps after fill in NAs")
```

Mean total number of steps taken per day.

```{r}
mean_step1 <- mean(total1$steps.per.day)
mean_step1
```

Median:

```{r}
median_step1 <- median(total1$steps.per.day)
median_step1
```

Do these values differ from the estimates from the first part of the assignment? Yes, they do.

Values without filling NAs:

Mean: 

```{r}
mean_step <- mean(total$steps.per.day)
mean_step
```

Median: 

```{r}
median_step <- median(total$steps.per.day)
median_step
```

Values after:

Mean: 

```{r}
mean_step1 <- mean(total1$steps.per.day)
mean_step1
```

Median: 

```{r}
median_step1 <- median(total1$steps.per.day)
median_step1
```

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
library(chron)
weekdays <- is.weekend(data1$date)

data1 <- data1 %>% mutate(weekday = weekdays)
data1$weekday <- as.character(data1$weekday)
data1$weekday <- gsub("FALSE", "weekday", data1$weekday)
data1$weekday <- gsub("TRUE", "weekend", data1$weekday)
data1$weekday <- as.factor(data1$weekday)

steps_by_interval_wday <- data1 %>% group_by(interval, weekday) %>% summarise(mean(steps))
names(steps_by_interval_wday)[3] <- "steps"
```

Create a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
library(ggplot2)
g <- ggplot(steps_by_interval_wday, aes(interval, steps))
g + 
        geom_line() + 
        facet_grid(weekday ~ .)
```